Evolving Dynamics: The U.S. Labor Force from 1972 to 2022

### 2024-03-18 paper/paper.qmd
updated discussion and appendix

## Potential improvements

@@

>>>>

## Potential Improvements

@@

## A Shift in Attitude

One major finding from the results is the change in attitudes towards women working. From the results, the attitudes towards women working have significantly increased over time. As seen in figure 6, the participants are asked is “better for men to work or women to tend home”. Around the year 1984, the two lines crossed indicating a changing point in attitude. The number of people who strongly agreed with this question has fallen under the number of people who strongly disagree. Another important time stamp to mention is in 2015, another interesting point of two lines. The number of people who answered “strongly disagree” even surpassed the number of people who agreed. From these two distinct indications in years 1984 and 2015, the data displays the change of attitude towards women in the workplace. Women are no longer expected to be the ones dedicated to tending home. Finally, looking at recent trends, both “disagree” and “strongly agree” have a sharp increase. The overall attitudes has shifted from unsupportive of women to work to more supportive attitudes. However, there is a big dip around 2021 in “disagree” and “strongly agree” and is also displayed in the other lines, this could be attributed to the overall decrease in data collection in recent years. More data would be needed to make further judgments on where the trend might be moving next.

## A Shift in Full Time Work and Higher Education

The trend of women participating in full-time work and education is also increasing. Upon taking a look at @fig-eduTojob, the number of graduate, undergraduate, and full-time lines is trending upwards. There are some significant jumps in the lines, however, this could be due to the data collection amount as these jumps are also observed in some other figures such as figure 8. With the attitudes steadily becoming more positive towards women in the workplace, the pursuit of higher education and full-time employment has also significantly increased. Something interesting observed is that the peak number of full-time workers is less than the most recent data. To gain insight into why this may be, additional research was done. In the Forbes article, it talks about the significant increase of women in higher education. With full-time working women decreasing, it could be attributed to them obtaining higher education instead of choosing to work now. Overall, this is a good trend as more women are pursuing higher education, and should allow them to obtain more skilled jobs in the future.

## Extension Effect of the Financial Crisis and Women in Work Place

Financial crises always have a devastating impact on society, how does it affect the women in the workforce? The financial crisis this section refers to is the 2008 housing bubble. There are also a few things that the data captured in the results section. In @fig-education, there is a sharp decrease in every line in the chart. The number of women that are participating in Full-time work, studying for graduation, or bachelor's all decreased. Upon doing additional research, in this paper. When an economic crisis occurs, the gender norms and biases we often forget about resurface. During the crisis, companies often cut costs by reducing labor. As a result, companies often decide to start with female workers because they are often seen as “secondary income” for the family. Something the article mentioned is loans taken out by women were impacted more, as lenders became more skeptical which leads to an impact on women-owned businesses. The number of graduate and bachelor women students also decreased, this could be because that family can no longer support the education, and decided that this would be a method of cutting living expenses. One example the paper mentioned is that women in India are expected to eat less as a result of reduced income to ensure that the rest of the family has enough food. As a result, women are often affected more by economic crises.

## Future Research Directions

### Supplementary Survey

To gain more insight into the problem, we have created our version of the survey. This survey is focused on getting answers from different perspectives on the issue. The questionnaire also incorporates considerations to reduce biases and also provide a bigger scale response other than just strongly disagree to strongly agree. Previous surveys do not provide enough insight into instances of women facing challenges in work place, the amount of questions that look into the attitudes toward women in the workplace is also very limited. As a result, this survey will aim to gain more insight with more detailed questions such as “Have you heard or witnessed a men colleague being paid less for the same position compared to women?”, the same question is also asked but for men. This will highlight the disproportionate amount of barriers and issues that women have to face compared to men and endure in the workplace. The data gathered from this survey can also be used to extend this current paper, allowing us to do a deeper analysis with the additional collected data.

Please see the @sec-appendix to view the full supplementary survey.

# Conclusion

With the data from the U.S. General Social Survey, the paper utilized it to gain further insight into the challenges women are enduring in the workplace. The time frame is slightly different in each figure but mostly focuses from around 1970 to 2022. To gain further insight, we looked at questions that could gain insight into attitudes towards women working and questions that could be affecting factors to women participating in the workforce. The data revealed that women overall have gained significant progress, but there are still many factors that need to be taken into consideration. Progress is indicated by a steady increase in positive attitudes towards women in the workplace, and also an increasing amount of women pursuing higher education. While these are good indicators of progress, there is still a lot to be uncovered. These questions do not provide for direct indication of the gender pay gap which is often the most brought up issue. More research and data gathering would need to be done to provide further insight into this tip-of-the-iceberg problem. We must continue to work to provide gender equality in the workplace, where there is still a lot of room for everyone to contribute to a more equal working environment. Let’s continue the momentum and ensure we continue working for an equal workforce.

# Appendix {#sec-appendix}

Here is the link to our supplementary survey:

<https://docs.google.com/forms/d/e/1FAIpQLSeF9oTRHWVcXK2nfLnddAI3mIWAVgcAJ3XnVthpz49frdeWFA/viewform>
